---
title: 'HCC: Predictive Model'
author: "Ph. Nesma Magdy Ahmed"
date: "22/04/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(Greg)
library(dplyr)
library(tidyr)
library(finalfit)
library(ggplot2)
library(moments)
library(e1071)
library(dummies)
library(dplyr)
library(tidyr)
library(finalfit)
library(dummies)
library(moments)
library(corrplot)
library(scatterplot3d)
library(ggplot2)
library(ggplot2)
library(caret)
library(GGally)
library(ggthemes)
library(broom)
library(dplyr)
library(bindrcpp)
library(caTools)
library(rattle)
library(RColorBrewer)
library(nnet)
library(rpart.plot)
library(dichromat)
library(Rmpfr)
library(digest)
library(convertr)
library(microbenchmark)
library(varhandle)
library(hablar)
library(magrittr)
library(survminer)
library(markovchain)
library(survival)
library(expss)
library(ggpubr)
library(tidyverse)
library(Rserve)
library(esquisse)
library(psych)
library(likert)
library(reshape)
library(usethis)
library(table1)
library(Hmisc)
library(psych)
library(MASS)
library(gtsummary)
library(ResourceSelection)
library(generalhoslem)
library(ordinal)
library(highcharter)
library(leaflet)
library(xlsx)
library(mlbench)
library(rcompanion)
library(caret)
library(performance)
library(foreign)
library(rms)
library(see)
library(rlang)
library(tidymodels)
library(janitor)
library(patchwork)
library(ROCR)
library(performance)
library(caret)
library(nricens)
```

## Data Cleaning:

```{r message=FALSE, warning=FALSE}

data <- read.csv(choose.files(), header = T, stringsAsFactors = T)
attach(data)

data[sapply(data[ ,1:23], is.numeric)] <- lapply(data[sapply(data[ ,1:23], is.numeric)], 
                                       as.factor)
dt <- read.csv(choose.files(), header = T, stringsAsFactors = T)
attach(dt)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

levels(data$Gender)[levels(data$Gender) == 0] <- "Female"
levels(data$Gender)[levels(data$Gender) == 1] <- "Male"


levels(data$Symptoms)[levels(data$Symptoms) == 0] <- "No Symptoms"
levels(data$Symptoms)[levels(data$Symptoms) == 1] <- "Symptoms"

levels(data$Alcohol)[levels(data$Alcohol) == 0] <- "Not Alcoholic"
levels(data$Alcohol)[levels(data$Alcohol) == 1] <- "Alcoholic"

levels(data$HBsAg)[levels(data$HBsAg) == 0] <- "No Hepatitis B Surface Antigen"
levels(data$HBsAg)[levels(data$HBsAg) == 1] <- "Hepatitis B Surface Antigen"

levels(data$HBeAg)[levels(data$HBeAg) == 0] <- "No Hepatitis B e Antigen"
levels(data$HBeAg)[levels(data$HBeAg) == 1] <- "Hepatitis B e Antigen"

levels(data$HBcAb)[levels(data$HBcAb) == 0] <- "No Hepatitis B Core Antibody"
levels(data$HBcAb)[levels(data$HBcAb) == 1] <- "Hepatitis B Core Antibody"

levels(data$HCVAb)[levels(data$HCVAb) == 0] <- "No Hepatitis C Virus Antibody"
levels(data$HCVAb)[levels(data$HCVAb) == 1] <- "Hepatitis C Virus Antibody"

levels(data$Cirrhosis)[levels(data$Cirrhosis) == 0] <- "No Cirrhosis"
levels(data$Cirrhosis)[levels(data$Cirrhosis) == 1] <- "Cirrhosis"

levels(data$Endemic)[levels(data$Endemic) == 0] <- "Not living in Endemic Countries"
levels(data$Endemic)[levels(data$Endemic) == 1] <- "Living in Endemic Countries"

levels(data$Smoking)[levels(data$Smoking) == 0] <- "Not Smoker"
levels(data$Smoking)[levels(data$Smoking) == 1] <- "Smoker"

levels(data$Diabetes)[levels(data$Diabetes) == 0] <- "Not Diabetic"
levels(data$Diabetes)[levels(data$Diabetes) == 1] <- "Diabetic"

levels(data$Obesity)[levels(data$Obesity) == 0] <- "Normal Weight"
levels(data$Obesity)[levels(data$Obesity) == 1] <- "Obese"

levels(data$Hemochro)[levels(data$Hemochro) == 0] <- "No Hemochromatosis"
levels(data$Hemochro)[levels(data$Hemochro) == 1] <- "Hemochromatosis"

levels(data$AHT)[levels(data$AHT) == 0] <- "No Arterial Hypertension"
levels(data$AHT)[levels(data$AHT) == 1] <- "Arterial Hypertension"

levels(data$CRI)[levels(data$CRI) == 0] <- "No Chronic Renal Insufficiency"
levels(data$CRI)[levels(data$CRI) == 1] <- "Chronic Renal Insufficiency"

levels(data$HIV)[levels(data$HIV) == 0] <- "No Human Immunodeficiency Virus (HIV)"
levels(data$HIV)[levels(data$HIV) == 1] <- "Human Immunodeficiency Virus (HIV)"

levels(data$NASH)[levels(data$NASH) == 0] <- "No Nonalcoholic Steatohepatitis"
levels(data$NASH)[levels(data$NASH) == 1] <- "Nonalcoholic Steatohepatitis"

levels(data$Varices)[levels(data$Varices) == 0] <- "No Esophageal Varices"
levels(data$Varices)[levels(data$Varices) == 1] <- "Esophageal Varices"

levels(data$Spleno)[levels(data$Spleno) == 0] <- "No Splenomegaly"
levels(data$Spleno)[levels(data$Spleno) == 1] <- "Splenomegaly"

levels(data$PHT)[levels(data$PHT) == 0] <- "No Portal Hypertension"
levels(data$PHT)[levels(data$PHT) == 1] <- "Portal Hypertension"

levels(data$PVT)[levels(data$PVT) == 0] <- "No Portal Vein Thrombosis"
levels(data$PVT)[levels(data$PVT) == 1] <- "Portal Vein Thrombosis"

levels(data$Metastasis)[levels(data$Metastasis) == 0] <- "No Liver Metastasis"
levels(data$Metastasis)[levels(data$Metastasis) == 1] <- "Liver Metastasis"

levels(data$Hallmark)[levels(data$Hallmark) == 0] <- "No Radiological Hallmark"
levels(data$Hallmark)[levels(data$Hallmark) == 1] <- "Radiological Hallmark"

levels(data$Class)[levels(data$Class) == 0] <- "Died"
levels(data$Class)[levels(data$Class) == 1] <- "Lived"

dt$Class1 <- dt$Class
levels(dt$Class1)[levels(dt$Class1) == 0] <- 1
levels(dt$Class1)[levels(dt$Class1) == 1] <- 0

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

data$Packs_year <- round(as.numeric(sub(",", ".", sub(".", " ,", data$Packs_year, fixed=TRUE), fixed=TRUE)),0)

data$INR <- round(as.numeric(sub(",", ".", sub(".", " ,", data$INR, fixed=TRUE), fixed=TRUE)),2)

data$AFP <- round(as.numeric(sub(",", ".", sub(".", " ,", data$AFP, fixed=TRUE), fixed=TRUE)),2)

data$Hemoglobin <- round(as.numeric(sub(",", ".", sub(".", " ,", data$Hemoglobin, fixed=TRUE), fixed=TRUE)),2)

data$MCV <- round(as.numeric(sub(",", ".", sub(".", " ,", data$MCV, fixed=TRUE), fixed=TRUE)),2)

data$Leucocytes <- round(as.numeric(sub(",", ".", sub(".", " ,", data$Leucocytes, fixed=TRUE), fixed=TRUE)),2)

data$Platelets <- round(as.numeric(sub(",", ".", sub(".", " ,", data$Platelets, fixed=TRUE), fixed=TRUE)),2)

data$Albumin <- round(as.numeric(sub(",", ".", sub(".", " ,", data$Albumin, fixed=TRUE), fixed=TRUE)),2)

data$Total_Bil <- round(as.numeric(sub(",", ".", sub(".", " ,", data$Total_Bil, fixed=TRUE), fixed=TRUE)),2)

data$ALT <- round(as.numeric(sub(",", ".", sub(".", " ,", data$ALT, fixed=TRUE), fixed=TRUE)),2)

data$AST <- round(as.numeric(sub(",", ".", sub(".", " ,", data$AST, fixed=TRUE), fixed=TRUE)),2)

data$GGT <- round(as.numeric(sub(",", ".", sub(".", " ,", data$GGT, fixed=TRUE), fixed=TRUE)),2)

data$ALP <- round(as.numeric(sub(",", ".", sub(".", " ,", data$ALP, fixed=TRUE), fixed=TRUE)),2)

data$TP <- round(as.numeric(sub(",", ".", sub(".", " ,", data$TP, fixed=TRUE), fixed=TRUE)),2)
data$Major_Dim <- dt$Major_Dim
data$Creatinine <- round(as.numeric(sub(",", ".", sub(".", " ,", data$Creatinine, fixed=TRUE), fixed=TRUE)),2)

data$Major_Dim <- round(as.numeric(sub(",", ".", sub(".", " ,", data$Major_Dim, fixed=TRUE), fixed=TRUE)),2)

data$Dir_Bil <- round(as.numeric(sub(",", ".", sub(".", " ,", data$Dir_Bil, fixed=TRUE), fixed=TRUE)),2)

data$Iron <- round(as.numeric(sub(",", ".", sub(".", " ,", data$Iron, fixed=TRUE), fixed=TRUE)),2)

data$Sat <- round(as.numeric(sub(",", ".", sub(".", " ,", data$Sat, fixed=TRUE), fixed=TRUE)),2)

data$Ferritin <- round(as.numeric(sub(",", ".", sub(".", " ,", data$Ferritin, fixed=TRUE), fixed=TRUE)),2)

data$Age <- as.numeric(data$Age)
data$Grams_day <- as.numeric(data$Grams_day)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

dt$Packs_year <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$Packs_year, fixed=TRUE), fixed=TRUE)),0)

dt$INR <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$INR, fixed=TRUE), fixed=TRUE)),2)

dt$AFP <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$AFP, fixed=TRUE), fixed=TRUE)),2)

dt$Hemoglobin <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$Hemoglobin, fixed=TRUE), fixed=TRUE)),2)

dt$MCV <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$MCV, fixed=TRUE), fixed=TRUE)),2)

dt$Leucocytes <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$Leucocytes, fixed=TRUE), fixed=TRUE)),2)

dt$Platelets <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$Platelets, fixed=TRUE), fixed=TRUE)),2)

dt$Albumin <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$Albumin, fixed=TRUE), fixed=TRUE)),2)

dt$Total_Bil <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$Total_Bil, fixed=TRUE), fixed=TRUE)),2)

dt$ALT <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$ALT, fixed=TRUE), fixed=TRUE)),2)

dt$AST <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$AST, fixed=TRUE), fixed=TRUE)),2)

dt$GGT <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$GGT, fixed=TRUE), fixed=TRUE)),2)

dt$ALP <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$ALP, fixed=TRUE), fixed=TRUE)),2)

dt$TP <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$TP, fixed=TRUE), fixed=TRUE)),2)

dt$Creatinine <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$Creatinine, fixed=TRUE), fixed=TRUE)),2)

dt$Major_Dim <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$Major_Dim, fixed=TRUE), fixed=TRUE)),2)

dt$Dir_Bil <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$Dir_Bil, fixed=TRUE), fixed=TRUE)),2)

dt$Iron <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$Iron, fixed=TRUE), fixed=TRUE)),2)

dt$Sat <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$Sat, fixed=TRUE), fixed=TRUE)),2)

dt$Ferritin <- round(as.numeric(sub(",", ".", sub(".", " ,", dt$Ferritin, fixed=TRUE), fixed=TRUE)),2)

dt$Class2 <- dt$Class1
dt$Class1 <- as.factor(dt$Class1)

```

## Demographic Characheristics:

```{r message=FALSE, warning=FALSE}

data %>% tbl_summary(by=Class) %>%
  bold_labels() %>%
  add_labelled_class()%>%
   modify_footnote(
    all_stat_cols() ~ "Median (IQR) or Frequency (%)"
  ) %>%
  add_p() %>%
  modify_header(label ~ "**Participants Characteristics**") %>%
  modify_caption("**Table 1. Patients' Characteristics (n = 204)**") %>%
  bold_labels() %>%
  bold_p()

```

## Splitting data into Test and Train:

By splitting the data set into random test and train data set will leading to potential colinearity issue.

```{r message=FALSE, warning=FALSE}

split1 <- sample(c(rep(0, 0.7 * nrow(data)), rep(1, 0.3 * nrow(dt))))

train <- dt[split1 == 0, ]
test <- dt[split1== 1, ] 

paste("train sample size: ", dim(train)[1])
paste("test sample size: ", dim(test)[1])

```


```{r fig.height=6, fig.width=13, message=FALSE, warning=FALSE}

cats.fit <- glm(Class1 ~ Symptoms + Endemic + PVT + Metastasis +
                  Age + PS + Encephalopathy + Ascites + INR + AFP +
                  Hemoglobin + Leucocytes + Albumin + Total_Bil + AST + GGT +
                  ALP + Creatinine + Major_Dim + Dir_Bil + Ferritin, data=train,
                family=binomial)
summary(cats.fit)

check_model(cats.fit)

```

After the model is created and fitted, this model is used for making predictions on the test data set values.

```{r message=FALSE, warning=FALSE}

pred <- predict(cats.fit, newdata = test, type = "response")
pred_bens <- as.factor(ifelse(pred>0.5,1,0))

```


```{r message=FALSE, warning=FALSE}

confusionMatrix(data = pred_bens, reference = test$Class1)

```

### Cutoff Values vs Accuracy:

```{r message=FALSE, warning=FALSE}

pred.rocr <- ROCR::prediction(pred, test$Class1)

eval <- ROCR::performance(pred.rocr, "acc")

plot(eval)

```

### Create AUC and ROC for test data:

ROC CURVE - ROC (Receiver Operator Characteristic Curve) can help in deciding the best threshold value. A ROC curve is plotted with FPR on the X-axis and TPR on the y-axis. A high threshold value gives - high specificity and low sensitivity A low threshold value gives - low specificity and high sensitivity.


```{r message=FALSE, warning=FALSE}
library(pROC)
test_roc <- roc(test$Class1 ~ pred, plot = TRUE, print.auc = TRUE)

auc(test_roc)

plot(test_roc,  
     print.cutoffs.at=-c(50,90,110,120), 
     cutoff.label.function=`-`,
     points.pch = 20, points.col = "darkblue", 
     text.adj=c(1.2,-0.5), legacy.axes = TRUE,
     print.auc = TRUE)

```

The model was verified in the external data set, AUC =0.824, indicating that the model has a good discrimination in the validation of the external data set.

##### As the random data splitting results in a potential colinearity issue.
##### Then, I will try the Bootstraping Method.

## Bootstraping:

```{r message=FALSE, warning=FALSE}

dt2 <- datadist(dt)

options(datadist = "dt2")

fit1 <- lrm(Class1~Symptoms + Endemic + PVT + Metastasis + Age + PS +
              Encephalopathy + Ascites + INR + AFP + Hemoglobin + Leucocytes +
              Albumin + Total_Bil + AST + GGT + ALP + Creatinine + Major_Dim +
              Dir_Bil + Ferritin, data = dt, x=T, y=T)

fit1

```

According to the calculated results, the C-Statistics is 0.886 in this data, which indicate a strong model.

```{r fig.height=13, fig.width=11, message=FALSE, warning=FALSE}

fun2 <- function(x)plogis(x-fit1$coefficients[1]+fit1$coefficients[2])
fun3 <- function(x)plogis(x-fit1$coefficients[1]+fit1$coefficients[3])

nom1 <- nomogram(fit1, fun = list('Prob Y>=1' = plogis, 'Prob Y>=2' = fun2),
                 fun.at = c(.01, .05, seq(.1, .9, by=1), .95, .99),
                 lp = F, funlabel = "Mortality Risk from HCC", minlength = 1,
                 abbrev = T)

plot(nom1)

```

```{r message=FALSE, warning=FALSE}

call <- calibrate(fit1, method = "boot", B=100)
plot(call, xlim=c(0, 1.0), ylim=c(0, 1.0))

```

it is obvious that the contribution of some predictors to the model are negligible, like, Encephalopathy, Albumin, Total Bilirubin, and Direct Bilirubin.

Exclude these variables that contributes less to the result from the regression model, then reconstruct model “fit2” and display the model parameters.

```{r message=FALSE, warning=FALSE}

fit2 <- lrm(Class1~Symptoms + Endemic + PVT + Metastasis + Age + PS + Ascites +
              INR + AFP + Hemoglobin + Leucocytes +  AST + GGT + ALP +
              Creatinine + Major_Dim + Ferritin, data = dt, x=T, y=T)

fit2

```

The C-Statistics = 0.884. The nomogram for the fit2 model become:

```{r fig.height=13, fig.width=11, message=FALSE, warning=FALSE}

fun2 <- function(x)plogis(x-fit1$coefficients[1]+fit1$coefficients[2])
fun3 <- function(x)plogis(x-fit1$coefficients[1]+fit1$coefficients[3])

nom2 <- nomogram(fit2, fun = list('Prob Y>=1' = plogis, 'Prob Y>=2' = fun2),
                 fun.at = c(.01, .05, seq(.1, .9, by=1), .95, .99),
                 lp = F, funlabel = "Mortality Risk from HCC",
                 minlength = 1, abbrev = T)

plot(nom2)

```

And the Calibration Model Become:

```{r message=FALSE, warning=FALSE}

call2 <- calibrate(fit2, method = "boot", B=100)
plot(call2, xlim=c(0, 1.0), ylim=c(0, 1.0))

```

### Calculation C-Statistics of Logistic Regression Model:

```{r message=FALSE, warning=FALSE}

dt$predvalue <- predict(fit1)

pred1 <- prediction(dt$predvalue, dt$Class1)

pref <- ROCR::performance(pred1, "tpr", "fpr")

plot(pref)
abline(0, 1, col = 3, lty = 2)

```

With Hmisc library:

```{r message=FALSE, warning=FALSE}
library(Hmisc)

somers2(dt$predvalue, dt$Class2)

```

Calculating the area under the ROC curve (AUC) is C-statistics = 0.8861015, which is consistent with the above calculation results.


### NRI calculation of dichotomous outcome:

Build a matrix out of a subset of the data containing Symptoms, Endemic, PVT, Metastasis, Age, PS, Encephalopathy, Ascites, INR, AFP, Hemoglobin, Leucocytes, Albumin, Total Bilirubin, AST, GGT, ALP, Creatinine, Major Dimension, Direct Bilirubin, and Ferritin.

```{r message=FALSE, warning=FALSE}

event <- dt$Class1

z.std <- as.matrix(subset(dt, select = c(Symptoms, Endemic, PVT, Metastasis,
                                         Age, PS, Encephalopathy, Ascites, INR,
                                         AFP, Hemoglobin, Leucocytes, Albumin,
                                         Total_Bil, AST, GGT, ALP, Creatinine,
                                         Major_Dim, Dir_Bil, Ferritin)))
  
```

Build a matrix out of a subset of data containing Symptoms, Endemic, PVT, Metastasis, Age, PS, Ascites, INR, AFP, Hemoglobin, Leucocytes, AST, GGT, ALP, Creatinine, Major Dimension, and Ferritin.

```{r message=FALSE, warning=FALSE}

z.new <- as.matrix(subset(dt, select = c(Symptoms, Endemic, PVT, Metastasis,
                                         Age, PS, Ascites, INR, AFP, Hemoglobin,
                                         Leucocytes, AST, GGT, ALP, Creatinine,
                                         Major_Dim, Ferritin)))
  
```

Construct two logistic regression model: mstd and mnew. Model “mnew” has one more variable “protime”. Calculation using “nricens” library requires x = TRUE, which means that output contains the matrix.

```{r message=FALSE, warning=FALSE}

mstd <- glm(event ~ Symptoms + Endemic + PVT + Metastasis + Age + PS +
              Encephalopathy + Ascites + INR + AFP + Hemoglobin + Leucocytes +
              Albumin + Total_Bil + AST + GGT + ALP + Creatinine + Major_Dim +
              Dir_Bil + Ferritin, binomial(logit),
            data.frame(dt$Class1, z.std), x=T)

mnew <- glm(event ~ Symptoms + Endemic + PVT + Metastasis + Age + PS + Ascites
            + INR + AFP + Hemoglobin + Leucocytes + AST + GGT + ALP +
              Creatinine + Major_Dim + Ferritin, binomial(logit),
            data.frame(dt$Class1, z.new), x=T)

```

##### Calculating the predicted risk of two models.

```{r message=FALSE, warning=FALSE}

nribin(mdl.std = mstd, mdl.new = mnew, cut = c(0.2, 0.4),
       niter = 100, updown = 'category')

```

##### Calculation of risk difference NRI using (‘event’, ‘z.std’, ‘z.new’)

```{r message=FALSE, warning=FALSE}

nribin(event = dt$Class2, z.std = z.std, z.new = z.new, cut = c(0.2, 0.4),
       niter = 100, updown = 'category')

```

##### Calculation of risk difference NRI using (‘mdl.std’, ‘mdl.new’), updown = ‘diff’:

```{r message=FALSE, warning=FALSE}

nribin(mdl.std = mstd, mdl.new = mnew, cut = 0.02, niter = 0, updown = 'diff')

```

##### Calculation of risk difference NRI using (‘event’, ‘z.std’, ‘z.new’), updown = ‘diff’:

```{r message=FALSE, warning=FALSE}

nribin(event = dt$Class2, z.std = z.std, z.new = z.new, cut = 0.02,
       niter = 100, updown = 'diff')

```

### IDI calculation of dichotomous outcome:

```{r message=FALSE, warning=FALSE}

library(PredictABEL)

p.std <- mstd$fitted.values
p.new <- mnew$fitted.values

dat_new <- cbind(dt, event)

reclassification(data = dat_new, cOutcome = 20, predrisk1 = p.std,
                 predrisk2 = p.new,
                 cutoff = c(0, 0.2, 0.4, 1))

```

IDI is 0.0039, indicating that new model improves 0.39% comparing to original model.

### Decision Curve Analysis for Binary Outcome:

We can view the data points on the complex model curve, where NB can also be changed to sNB, indicating a standardized prevalence.

```{r message=FALSE, warning=FALSE}

library(rmda)

complex <- decision_curve(Class2~Symptoms + Endemic + PVT + Metastasis +
                            Age + PS + Encephalopathy + Ascites + INR + AFP +
                            Hemoglobin + Leucocytes + Albumin + Total_Bil + AST
                          + GGT + ALP + Creatinine + Major_Dim + Dir_Bil +
                            Ferritin, data = dt, 
                          family = binomial(link = 'logit'),
                          thresholds = seq(0, 1, by = 0.01),
                          confidence.intervals = 0.95,
                          study.design = "cohort",
                          population.prevalence = 0.3)

summary(complex, measure = 'NB')

plot_clinical_impact(complex, population.size = 1000, cost.benefit.axis = T,
                     n.cost.benefits = 8, col = c('red', 'blue'),
                     confidence.intervals = T,
                     ylim = c(0, 1000),
                     legend.position = 'topright')
```

The red curve (number of high risk) indicates the number of people classified as positive (high risk) by the complex model at each threshold probability; the blue curve [(number high) risk with outcome] is the number of true positives for each threshold probability.

#### External validation of Logistic regression model:

```{r message=FALSE, warning=FALSE}

mod <- glm(Class2 ~ Symptoms + Endemic + PVT + Metastasis + Age + PS +
             Encephalopathy + Ascites + INR + AFP + Hemoglobin + Leucocytes +
             Albumin + Total_Bil + AST + GGT + ALP + Creatinine + Major_Dim +
             Dir_Bil + Ferritin, data=dt, family=binomial)

h1 <- hoslem.test(mod$y, fitted(mod), g=10)
h1

```

P value is 0.3026, It cannot be considered that the model does not fit well yet.

```{r message=FALSE, warning=FALSE}

cbind(h1$observed, h1$expected)

```

Generate Hosmer-Lemeshow contingency table. Among them, y0 is the number of events that didn’t occur; y1 is the number of events that occurred; yhat0 represents the probability of events that will not happen predicted by the model; yhat1 represents the probability of events that will happen predicted by the model.

```{r message=FALSE, warning=FALSE}

pre <- predict(mod, dt, type = c("response"))

h2 <- hoslem.test(dt$Class2, pre, g=10)
h2

```
P=0.3026 > 0.05. So, it cannot be considered that the model fitting is poor, suggesting that the model performs well in the new data set. If P < 0.05, the model is poorly fitted.


```{r message=FALSE, warning=FALSE}

library(PredictABEL) 
plotCalibration(data = dt, 
                cOutcome = 2,
                predRisk = pre,
                rangeaxis = c(0, 1))

```


```{r message=FALSE, warning=FALSE}

library(PredictABEL)
library(pROC)
plotCalibration(data = dt, 
                cOutcome = 2,
                predRisk = pre,
                rangeaxis = c(0, 1))

```

#### Discrimination evaluation:

```{r message=FALSE, warning=FALSE}

library(pROC)
pr <- predict(mod, type = c("response"))

roc_curve <- pROC::roc(dt$Class2 ~ pr)


auc(roc_curve)


plot(roc_curve,  
     print.cutoffs.at=-c(50,90,110,120), 
     cutoff.label.function=`-`,
     points.pch = 20, points.col = "darkblue", 
     text.adj=c(1.2,-0.5), legacy.axes = TRUE,
     print.auc = TRUE)

```

The model was verified in the external data set, AUC =0.8861, indicating that the model has a good discrimination in the validation of the external data set.




































